A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-11-25 01_Introduction.Rmd
style vs. function

Archaeology has a special perspective to offer for Cultural evolution research: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018]. Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of live. We rightfully expect humans to learn handcraft as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them. The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].

In 1978 Robert Dunnell formulated in a contribution a previously discussed and in the following intensive to a distinction that has been taken up: In archaeological contexts and over long periods of time, ideas would have to be and behaviour patterns into two categories function (functions) and form (style). Functional Ideas on how the knowledge of the use of a cereal form, the competence of certain cereal Tools or weapons to manufacture or social recognition of child abundance are directly relevant to selection. This is where the biological evolution, the survival rate in crisis situations, or of demographic development, to favor certain ideas about their carriers. This Ideas are opposed by many others: Newtral traits that are not directly relevant for selection. such as ceramic decorations, traditional costume fashion or Jewelry shape. Survival or reproductive ability of a population are usually not affected by the nature of the geometric patterns on their storage vessels dependent. However, these characteristics pass through a partly breathtakingly fast development and allow archaeologists insights into relative chronology and social structure. According to Dunnell, in the context of the of Evolutionary Archaeology despite its genetically subordinate clout deserve special relevance - and a dedicated set of methods: Dunnell postulates, that functional ideas, taking into account the quality of their adaptation could be modelled, while stylistic Ideas essentially random, stochastic processes (Neutral Transmission) would follow.

>>>>

Archaeology has a special perspective to offer for Cultural evolution research: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018]. Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of live. We rightfully expect humans to learn handcraft as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them. The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002]. In this context a distinction of culture in function and style became an important question for *Evolutionary archaeology*, explicitly elaborated already by [@dunnell1978style]. Some types of human behaviour obviously increase wellbeing and simplify survival, like certain hunting techniques, superior tools or sowing more fruitful crops. Without resorting to the inappropriate simplifications of Evolutionism their diffusion can be expected to increase over time. On the other hand there are behaviours that are seemingly not related to survival and reproductive capacity, like e.g. ceramic decorations, traditional costume fashion or jewelry design.

However, these characteristics pass through a partly breathtakingly fast development and allow archaeologists insights into relative chronology and social structure. According to Dunnell, in the context of the of Evolutionary Archaeology despite its genetically subordinate clout deserve special relevance - and a dedicated set of methods: Dunnell postulates, that functional ideas, taking into account the quality of their adaptation could be modelled, while stylistic Ideas essentially random, stochastic processes (Neutral Transmission) would follow.
